Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-09-26 decorrelation/decorrelation.tex
adding small blurb about least squares, citation still needed

This can be solved with a constrained minimization algorithm, such as Sequential Least Squares Programming\cite{wright1999numerical}.

@@

Using Equation \ref{eq:minimization}, we then generated the optimal (IS THIS CONVEX?) weights $\omega_i$. To benchmark the quality of the new balanced channel, we then performed six randomized benchmarking experiments: one for each over- and under-calibrated pulse, one for the calibrated pulse, and one for the balanced channel. We used $N=1000$ shots per experiment and $K=10$ sequences per sequence length, for sequence lengths up to $L=64$\cite{Magesan2011}. In each case, our Clifford operations were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses. In our implementation, these gates are implemented using the same pulse envelope definitions and control electronics, phase shifted by $\frac{\pi}{2}$ radians, and are therefore subject to identical miscalibration errors. The results are shown in Figure \ref{fig:rb} for sequence lengths $L=64$. In this experiment, fitting the randomized benchmarking data with scipy's curve\_fit using the Trust Region Reflective method reports one-qubit gate fidelities of $99.4\%$ for the calibrated pulse, $98.8\%$ for Pulse1, $99.4\%$ for Pulse2, $98.9\%$ for Pulse3, $98.6\%$ for Pulse4, and $99.0\%$ for the Randomized pulse. As one can see from looking at Figure \ref{fig:rb}, despite all controls performing comparably to the Calibrated control, only the Randomized control has similarly tight error bars. In the cases of each miscalibrated pulse, the error bars extend significantly, in often asymmetric ways, making usage of chi-square statistics, and errors on the fit less straightforward.

@@

In these situations, one approach would be to use \textit{in situ} optimal control techniques \cite{Wu2018, Kelly2014, Ferrie2015} to generate candidate controls, and then use an optimizer like Nealder-Mead to perform the minimization. While performing optimization this way would still be slow (requiring full process tomography, in general, to reconstruct the process matrix of a candidate balanced channel), one could intead select pulse sequences that are sensitive to known parameters of interest in the Hamiltonian and just run the minimization with respect to those terms. (e.g. if we suspect there is unwanted $Z\otimes Z$ crosstalk between qubits.)

>>>>

In the numerics section below, this was solved with a constrained minimization algorithm (Sequential Least Squares Programming\cite{wright1999numerical}). However, it can be solved more efficiently as a least squares problem. In particular, consider the $M^2-M$ off diagonal elements $\mathcal{E}_i$. Vectorize their squares, and concatenate the resulting vectors, for each $\mathcal{E}_i$. Then we just need to find $\omega$ that produces the smallest $\vec{b}$ in the following underconstrainted least squares problem:
\begin{center}
$A\omega= \vec{b}$
\end{center}
where:
\begin{center}

A = $\begin{pmatrix}
\mathcal{E}_{0_{10}} & \mathcal{E}_{1_{10}} & \cdots & \mathcal{E}_{N-1_{10}}\\
\mathcal{E}_{0_{20}} & \mathcal{E}_{1_{20}} & \cdots &\mathcal{E}_{N-1_{20}}\\
\vdots & \vdots & \ddots & \\
\mathcal{E}_{0_{M-1 M}} &\mathcal{E}_{1_{M-1 M}} & & \mathcal{E}_{N-1_{M-1 M}}
\end{pmatrix}
$
\end{center}
In particular, we want the solution $\omega$ such that $||\vec{b}||_2^2$ is minimized. This is readily given by the solution to a constrained least squares minimization problem as $\frac{1}{N}(\tilde{A^{-1}}_{AVG})^{-1}\tilde{A^{-1}}_{i_{AVG}}$.\needcite\url{http://eeweb.poly.edu/iselesni/lecture_notes/least_squares/least_squares_SP.pdf}

@@

Using Equation \ref{eq:minimization}, we then generated the optimal weighting $\omega_i$. To benchmark the quality of the new balanced channel, we then performed six randomized benchmarking experiments: one for each over- and under-calibrated pulse, one for the calibrated pulse, and one for the balanced channel. We used $N=1000$ shots per experiment and $K=10$ sequences per sequence length, for sequence lengths up to $L=64$\cite{Magesan2011}. In each case, our Clifford operations were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses. In our implementation, these gates are implemented using the same pulse envelope definitions and control electronics, phase shifted by $\frac{\pi}{2}$ radians, and are therefore subject to identical miscalibration errors. The results are shown in Figure \ref{fig:rb} for sequence lengths $L=64$. In this experiment, fitting the randomized benchmarking data with scipy's curve\_fit using the Trust Region Reflective method reports one-qubit gate fidelities of $99.4\%$ for the calibrated pulse, $98.8\%$ for Pulse1, $99.4\%$ for Pulse2, $98.9\%$ for Pulse3, $98.6\%$ for Pulse4, and $99.0\%$ for the Randomized pulse. As one can see from looking at Figure \ref{fig:rb}, despite all controls performing comparably to the Calibrated control, only the Randomized control has similarly tight error bars. In the cases of each miscalibrated pulse, the error bars extend significantly, in often asymmetric ways, making usage of chi-square statistics, and errors on the fit less straightforward.

@@

In these situations, one approach would be to use \textit{in-situ} optimal control techniques \cite{Wu2018, Kelly2014, Ferrie2015} to generate candidate controls, and then use an optimizer like Nealder-Mead to perform the minimization. While performing optimization this way would still be slow (requiring full process tomography, in general, to reconstruct the process matrix of a candidate balanced channel), one could intead select pulse sequences that are sensitive to known parameters of interest in the Hamiltonian and just run the minimization with respect to those terms. (e.g. if we suspect there is unwanted $Z\otimes Z$ crosstalk between qubits.)
